---
title: "Covid-19"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(htmltools)
library(htmlwidgets)
library(ggplot2)
library(dplyr)
library(DT)
library(leaflet)
library(rgdal)
library(RColorBrewer)
library(plotly)
library(zoo)
library(scales)
library(shiny)
library(rsconnect)
library(gapminder)
library(sp)
library(stats)
```
 

```{r, include=FALSE}
# data <- read.csv('https://covid.ourworldindata.org/data/owid-covid-data.csv', header = T)
data <- read.csv("owid-covid-data.csv", header = T)
```

```{r, echo = F}
cum_dat <- data %>%
  filter(as.Date(date) == max(as.Date(date))) %>%
    select(iso_code,
           Country = location,
           continent,
           Population = population, 
           Cases = total_cases,
           Deaths = total_deaths,
           population_density,
           median_age,
           gdp_per_capita,
           extreme_poverty,
           cvd_death_rate,
           diabetes_prevalence,
           life_expectancy) %>%
  mutate(Mortality = round(Deaths/Cases*100, 3)) %>%
  arrange(desc(Cases))

```

Global Cases
===================================== 

Row
-------------------------------------
### Global
```{r}
valueBox("Covid-19 Analysis")
```

### Confirmed Cases
```{r}
valueBox(cum_dat$Cases[cum_dat$Country == "World"])
```

### Deaths
```{r}
valueBox(cum_dat$Deaths[cum_dat$Country == "World"])
```

### Mortality Rate
```{r}
valueBox(paste(round(cum_dat$Deaths[cum_dat$Country == "World"]/cum_dat$Cases[cum_dat$Country == "World"]*100, 2), "%", sep = ""))
```

Row {.tabset}
-------------------------------------
### New Daily Infections
```{r}
#Creating a Subsetted Dataframe
world_dat <- data %>% 
  filter(location == "World") %>% 
  mutate(date = as.Date(date),
         new_cases_SMA7 = rollmean(new_cases, k = 7, align = "center", fill = NA))

#Creating Plot
fig <- plot_ly(world_dat, 
               x = ~date, 
               y = ~new_cases, 
               type = "bar",
               name = "New Cases"
               )

#Adding Moving Averages
fig <- fig %>% add_trace(y = ~new_cases_SMA7,
                         name = "7-Day Moving Average", 
                         type = "scatter", 
                         mode = "lines", 
                         line = list(color = 'rgb(250, 162, 0)'))
fig <- fig %>% layout(yaxis = list(title = "New Cases"), 
                      xaxis = list(title = "Date")
  
)
fig
```

### Suseptible-Infected-Deaths
```{r}
fig <- plot_ly(world_dat, x = ~date, y = ~population-total_cases, name = "Suseptible", type = 'scatter', mode = 'lines')
fig <- fig %>% add_trace(y = ~total_cases, name = "Infections", type = 'scatter', mode = 'lines')
fig <- fig %>% add_trace(y = ~total_deaths, name = "Deaths", type = 'scatter', mode = 'lines')
fig <- fig %>% layout(xaxis = list(title = "Date"), yaxis = list(title = "Population"))
fig
```

  
### Important Information
  This model is not an accurate representation of the SIR model because there is insufficient data tracking individuals who have recovered from COVID-19: thus, infections is not an accurate representation of current infectious and wil contain both infected and infectious individuals. Where the model calls for $S$ = the number of susceptible individuals, $I$ = the number of infectious individuals, $R$ = individuals who have recovered or have deceased.
  
  In addition, the removed has been replaced with death due to the same reason regarding the data. For more information on the SIR model read the section named additional information.
  
  Please note that this model does not account for changing populations.



Row
-------------------------------------
### Cases by Country

```{r}
# Text Data
cum_dat$hover <- with(cum_dat, paste(Country, '<br>', 
                                     "Cases", Cases, '<br>',
                                     "Deaths", Deaths, 
                                     "<br>", "Mortality", paste(Mortality, "%", sep = "")))
## Plotting Cases accross the world
#Boarder Colors
l <- list(color = toRGB("grey"), width = 0.5)

# specify map projection/options
g <- list(
  showframe = FALSE,
  showcoastlines = FALSE,
  projection = list(type = 'Mercator')
)

fig <- plot_geo(cum_dat)
fig <- fig %>% add_trace(
    z = ~Cases, color = ~Cases, colors = 'Blues', 
    text = ~hover, locations = ~iso_code, marker = list(line = l)
  )
fig <- fig %>% colorbar(title = 'Cases')
fig <- fig %>% layout(autosize = T, 
                      # width = 500, 
                      # height = 500,
                       geo = g)
fig
```


### Table of Cumulative Confirmed Cases, Deaths, and Tests
```{r}
datatable(
  data %>% group_by(location) %>%
  filter(as.Date(date) == max(as.Date(date)), location != "World") %>%
    select(Country = location, 
           Population = population, 
           Cases = total_cases, 
           "Cases Per Million" = total_cases_per_million,
           Deaths = total_deaths, 
           "Deaths Per Million" = total_deaths_per_million) %>%
  mutate("PopulationInfected (%)" = round((Cases/Population)*100, 3),
         "Mortality (%)" = round(Deaths/Cases*100, 3)) %>%
  arrange(desc(Cases))
)
```


Total Cases By Country
===================================== 

Column {.sidebar}
-------------------------------------
```{r}
# Dataset
by_days <- data %>% group_by(location) %>% filter(total_cases >= 1, location != "World") %>%
  mutate(days = as.numeric(-rev(difftime(as.Date(date), max(as.Date(date)), units = "days"))),
         new_cases_SMA7 = rollmean(new_cases, k = 7, align = "center", fill = NA)) %>%
  select(location, total_cases, days, population, new_cases, continent, new_cases_SMA7)


selectInput("c_1", "Select a country", sort(unique(by_days$location)), selected = 'United States')
selectInput("c_2", "Add a country", sort(unique(by_days$location)), selected = 'Brazil')
selectInput("c_3", "Add a country", sort(unique(by_days$location)), selected = 'India')
selectInput("c_4", "Add a country", sort(unique(by_days$location)), selected = 'Italy')
selectInput("c_5", "Add a country", sort(unique(by_days$location)), selected = 'Philippines')
selectInput("c_6", "Add a country", sort(unique(by_days$location)), selected = 'South Africa')
selectInput("c_7", "Add a country", sort(unique(by_days$location)), selected = 'China')
selectInput("c_8", "Add a country", sort(unique(by_days$location)), selected = 'Japan')
selectInput("c_9", "Add a country", sort(unique(by_days$location)), selected = 'Germany')
selectInput("c_10", "Add a country", sort(unique(by_days$location)), selected = 'Lebanon')
```


Column 
-------------------------------------
### Growth Comparison by Population Proportion

```{r}
plotly::renderPlotly({
fig <- plot_ly(by_days[which(by_days$location %in% c(input$c_1,
                                             input$c_2,
                                             input$c_3,
                                             input$c_4,
                                             input$c_5,
                                             input$c_6,
                                             input$c_7,
                                             input$c_8,
                                             input$c_9,
                                             input$c_10)),], x = ~days, y = ~total_cases/population*100, color = ~location, type = "scatter", mode = "lines")
fig <- layout(fig, yaxis = list(title = "Percent (%) of Total Population Infected"), xaxis = list(title = "Days Since First Case"))
fig
})

```

Row
-------------------------------------
### Growth Comparison by Infection Size

```{r}
plotly::renderPlotly({
fig <- plot_ly(by_days[which(by_days$location %in% c(input$c_1,
                                             input$c_2,
                                             input$c_3,
                                             input$c_4,
                                             input$c_5,
                                             input$c_6,
                                             input$c_7,
                                             input$c_8,
                                             input$c_9,
                                             input$c_10)),], x = ~days, y = ~total_cases, color = ~location, type = "scatter", mode = "lines")
fig <- layout(fig, yaxis = list(title = "Total Cases (log)",type = "log"), xaxis = list(title = "Days Since First Case"))
fig})


```


Growth By Country
===================================== 

Column {.sidebar}
-------------------------------------

```{r}
selectInput("country1", "Select a country", sort(unique(by_days$location)), selected = 'United States')
selectInput("country2", "Add a country", sort(unique(by_days$location)), selected = 'Brazil')
selectInput("country3", "Add a country", sort(unique(by_days$location)), selected = 'India')
```

Column
-------------------------------------
```{r}
plotly::renderPlotly({
  # Input 1
 fig <- plot_ly(by_days[which(by_days$location %in% c(input$country1,input$country2,input$country3)),], 
                x = ~days, 
                y = ~new_cases, 
                color = ~location,
                type = "bar")
 
 fig <- fig %>% add_trace(y = ~new_cases_SMA7, 
                         color = ~location,
                         name = "7-Day Moving Average", 
                         type = "scatter", 
                         mode = "lines")
 fig <- fig %>% layout(xaxis = list(title = "Days Since First Case"), 
                       yaxis = list(title = "Daily New Cases"))
 fig
})


```


Explanatory Factors
===================================== 

Column {.sidebar}
-------------------------------------
```{r}
selectInput("Independent", "Select the variable for the X-axis (Explanatory Variable)", choices = c("Population Density", "Median Age", "GDP per Capita", "Extreme Poverty", "CVD Death Rate", "Diabetes Prevalence", "Life Expectancy"), selected = "GDP per Capita")

selectInput("Dependent", "Select the variable for the Y-axis (Dependent Variable)",  choices = c("Cases", "Deaths", "Mortality"), selected = "Mortality")

selectInput("y_scale", "Select the scale of the y-axis", choices = c("linear", "log"), selected = "linear")
selectInput("x_scale", "Select the scale of the x-axis", choices = c("linear", "log"), selected = "linear")
```

Column
-------------------------------------

```{r}
indep <- reactive({
if ( "Population Density" %in% input$Independent) return(cum_dat[which(cum_dat$Country != "World"),]$population_density)
if ( "Median Age" %in% input$Independent) return(cum_dat[which(cum_dat$Country != "World"),]$median_age)
if ( "GDP per Capita" %in% input$Independent) return(cum_dat[which(cum_dat$Country != "World"),]$gdp_per_capita)
if ( "Extreme Poverty" %in% input$Independent) return(cum_dat[which(cum_dat$Country != "World"),]$extreme_poverty)
if ( "CVD Death Rate" %in% input$Independent) return(cum_dat[which(cum_dat$Country != "World"),]$cvd_death_rate)
if ( "Diabetes Prevalence" %in% input$Independent) return(cum_dat[which(cum_dat$Country != "World"),]$diabetes_prevalence)
if ( "Life Expectancy" %in% input$Independent) return(cum_dat[which(cum_dat$Country != "World"),]$life_expectancy)
  })

dep <- reactive({
if ( "Cases" %in% input$Dependent) return(cum_dat[which(cum_dat$Country != "World"),]$Cases)
if ( "Deaths" %in% input$Dependent) return(cum_dat[which(cum_dat$Country != "World"),]$Deaths)
if ( "Mortality" %in% input$Dependent) return(cum_dat[which(cum_dat$Country != "World"),]$Mortality)
  })

renderPlotly({
 fig <- cum_dat[which(cum_dat$Country != "World"),] %>% plot_ly(x = ~indep(), y = ~dep(), color = ~continent, text = ~Country, hoverinfo = "text", type = 'scatter', mode = 'markers', tendline = "ols")
 fig <- fig %>% layout(yaxis = list(title = input$Dependent,type = input$y_scale), xaxis = list(title = input$Independent, type = input$x_scale))
 fig
})
```


Additional Information
=====================================

  
The dataset was obtained by Our World in Data and can be obtained here:   https://github.com/owid/covid-19-data/tree/master/public/data.  
  
The purpose of this project was to implement newly developed skills in R with Flexdashboard and the shiny app; in addition, it was used to explore some of the mathematical models in epidemiology.  
  
Hopefully this app has provided useful information and that the information is presented in a way which is easily understood. More information will be added into this section regarding the mathematical models surrounding diseases and infections including a possible simulation done in NetLogo or R.  
  
For more information on the disease itself:  
  https://www.who.int/emergencies/diseases/novel-coronavirus-2019  
  https://www.cdc.gov/
  
Github:  https://github.com/CuddlyCuddle  
  
  
